Highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2020-01-17 write_up/mixfishsim_ecolmod_revisions.tex
Edited to M&M

\usepackage[draft]{changes}

@@

\address[1]{Galway-Mayo Institute of Technology (GMIT), Dublin Road, Galway,
	Ireland} 
\address[2]{Centre for Environment, Fisheries and Aquaculture Science (Cefas),
	Pakefield Road, Lowestoft, UK}
\address[3]{Sorbonne Université, Faculty of Sciences, 4 Place Jussieu, 75005
	Paris, France}
\address[4]{Wageningen Marine Research, Haringkade 1 1976 CP IJmuiden,
	Netherlands}
\address[5]{Aquaculture and Fisheries Group, Wageningen University \& Research,
	Zodiac Building 122, De Elst 1, 6708 WD Wageningen, the Netherlands}

@@

%%
To understand how data resolution impacts inference on mixed fisheries
interactions we developed a highly resolved spatiotemporal discrete-event
simulation model ('MixFishSim`) incorporating: i) delay-difference population
dynamics, ii) population movement using Gaussian Random Fields to simulate
patchy, heterogeneously distributed \added{and moving fish} populations, and
iii) fishery dynamics for multiple fleet characteristics based on population
targeting under an explore-exploit strategy.\deleted{ This is implemented via a mix of
correlated random walk movement (for exploration) and learned behaviour (for
exploitation) phases of the fisheries. }\\ 
%%
\deleted{Fifty years of sub-daily fishing activity was simulated and used to
	draw inference on the underlying community structures. W}\added{We
	applied the framework to draw}\deleted{compared} inference\deleted{s}
\added{on community structure} \added{when using data generated from
}\deleted{based on}: commercial catch, a simulated fixed-site sampling survey
design and the true \added{(simulated)} underlying populations.
\deleted{W}\added{In doing so w}e \deleted{i)}establish the potential
limitations of fishery-dependent data in providing a robust picture of
spatiotemporal distributions\deleted{; and then ii) simulated an area closure
	based on areas defined from the different data sources at a range of
	temporal and spatial resolutions.}\deleted{Our framework allows users
	to explore the assumptions in modelling observational data and evaluate
	the underlying dynamics of such approaches at a fine spatial and
	temporal scale. In application to a mixed fishery exploiting four
	different populations w} \added{W}e found different spatial patterns
were evident and that the effectiveness of the spatial closure reduced when
data were aggregated across larger spatial areas. \deleted{; and then
	ii)}\added{We then simulated an area closure based on areas defined
	from the different data sources at a range of temporal and spatial
	resolutions.} Aggregation across time periods has less of a negative
impact on the closure success than over space.\deleted{ and w}\added{W}hile not
as effective as when based on on the true population, closures based on high
catch rates observed in commercial data were still able to reduce fishing on a
protected species. \\

\added{Our framework allows users to explore the assumptions in modelling
	observational data and evaluate the underlying dynamics of such
	approaches at a fine spatial and temporal scale. }\added{From our
	application w}\deleted{W}e conclude that commercial data, while
containing bias, provide a useful tool for managing catches in mixed fisheries
if applied at the correct spatiotemporal scale. \\
%%
\end{abstract}

@@

Identifying appropriate spatial scales for fisheries closures has been a
highlighted as crucial to their success \citep{Costello2010, Dunn2016}.
Inference on fisheries spatial dynamics is hampered where spatial information
is coarse due to low resolution reporting of fisheries catch which is
aggregated across larger gridded areas \citep{Branch2005}. Further, if data
does not allow identification of spatial features it may lead to poorly sited
closures which are ineffectual or have unintended consequences. For example,
increased benthic impact on previously unexploited areas from the cod closure
in the North Sea were observed without the intended effect of reducing cod
exploitation \citep{Rijnsdorp2001,Dinmore2003}).\\

More refined spatiotemporal information has since become available through the
combination of logbook and Vessel Monitoring System (VMS) data \citep{Lee2010,
	Bastardie2010, Gerritsen2012, Mateo2016} and more real-time spatial
management has been possible \citep[e.g.][]{Holmes2011}.  However, fishers
establish favoured fishing grounds through an explore-exploit strategy
\citep{Rijnsdorp2011, Bailey2018} where they search for areas with high catches
and then use experience to return to areas where they've experienced high catch
in the past. This leads to an inherently biased sampling where target species
are over-represented in the catch as fishers exploit areas of high abundance.
There is a need to understand the influence of these biases on any spatial
management measures which are implemented based on inference from commercial
landings or catch data. \\

@@

\begin{enumerate}
	\item Do different source of sampling-derived fisheries data
		reflect\deleted{s} the underlying community structure? 	
	\item How does data aggregation and data source impact on the success
		of spatial fisheries management measures?
\end{enumerate}

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time, using a Gaussian spatial
process to define suitable habitat for different populations. As the precise
locations of the fish are known directly rather than inferred from sampling
or commercial catch, we can use the population model to validate how inference
from fisheries-dependent and fisheries independent sampling relates to the real
community structure in a way we could not with real data. We ii) compare, at
different spatial and temporal aggregations, the real (simulated) population
distributions to samples from fisheries-dependent and fisheries independent
catches to test if these are a true reflection of the relative density of the
populations. We then iii) simulate a fishery closure to protect a species based
on different spatial and temporal data aggregations. \\

@@

Fishing locations were chosen based on random search and, with increasing
proportion as time progressed, experience of profitable catches built up in the
same month from previous years and from the previous trip.  'Profitable' in
this context was defined as the locations where the top 70 \% of expected
profit would be found given revenue from previous trips and cost of movement to
the new fishing location. This probability was based on a logistic sigmoid
function with a lower asymptote of 0 and upper asymptote of 0.95, and a growth
rate that ensures the upper asymptote (where decisions are mainly based on past
knowledge) is reached approximately halfway through the simulation.  \\

@@

\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_space_Rev2}
	\caption{[Colour] Data aggregation at different spatial resolutions
		over a ten year period. \added{The figure shows catch
			composition at each spatial unit represented by a
			square pie chart of the four populations. The area of
			each colour is proportionate to the weight of each
			population caught in that unit. Figure produced using
			the R package 'mapplots' (\cite{Gerritsen2014}.}
	}
	\label{fig:1}
\end{figure}

>>>>

\usepackage[draft]{changes}%\usepackage[draft]{changes}

@@

\address[1]{Marine and Freshwater Research Centre, Galway-Mayo Institute of Technology (GMIT), Dublin Road, Galway,
	Ireland} 
\address[2]{Centre for Environment, Fisheries and Aquaculture Science (Cefas),
	Pakefield Road, Lowestoft, UK}
\address[3]{Sorbonne Université, Faculty of Sciences, 4 Place Jussieu, 75005
	Paris, France}
\address[4]{Wageningen Marine Research, Haringkade 1 1976 CP IJmuiden,
	Netherlands}
\address[5]{Aquaculture and Fisheries Group, Wageningen University \& Research,
	Zodiac Building 122, De Elst 1, 6708 WD Wageningen, the Netherlands}

@@

%%
To understand how data resolution impacts inference on mixed fisheries
interactions we developed a highly resolved spatiotemporal discrete-event
simulation model (`MixFishSim')\textcolor{blue}{[CM: I think \emph{MixFishSim} is more effective]} incorporating: i) delay-difference population
dynamics, ii) population movement using Gaussian Random Fields to simulate
patchy, heterogeneously distributed \added{and moving fish} populations, and
iii) fishery dynamics for multiple fleet characteristics based on population
targeting under an explore-exploit strategy.\deleted{ This is implemented via a mix of
correlated random walk movement (for exploration) and learned behaviour (for
exploitation) phases of the fisheries. }\\ 
%%
\deleted{Fifty years of sub-daily fishing activity was simulated and used to
	draw inference on the underlying community structures. W}\added{We
	applied `MixFishSim' to\deleted{draw}}\deleted{compared} infer\deleted{ence}\deleted{s}
\added{\deleted{on}community structure} \added{when using data generated from:}
\deleted{based on}commercial catch, a\deleted{simulated} fixed-site sampling survey
design and the true \added{(simulated)} underlying populations.
\deleted{W}\deleted{\added{In doing so w}}\added{W}e \added{thereby} \deleted{i)}establish the potential
limitations of fishery-dependent data in providing a robust \replaced{characterisation}{picture} of
spatiotemporal distributions.\deleted{; and then ii) simulated an area closure
	based on areas defined from the different data sources at a range of
	temporal and spatial resolutions.}\deleted{Our framework allows users
	to explore the assumptions in modelling observational data and evaluate
	the underlying dynamics of such approaches at a fine spatial and
	temporal scale. In application to a mixed fishery exploiting four
	different populations w} \deleted{\added{W}e found d}\added{D}ifferent spatial patterns
were evident and \deleted{that }the effectiveness of the spatial closure reduced when
data were aggregated across larger spatial areas. \deleted{; and then
	ii)}\deleted{\added{We then simulated an area closure based on areas defined
	from the different data sources at a range of temporal and spatial
	resolutions.}} \added{A simulated area closure showed that} \replaced{a}{A}ggregation across time periods has less of a negative
impact on the closure success than \added{aggregation} over space. \deleted{ and w}\added{W}hile not
as effective as when based on on the true population, closures based on high
catch rates observed in commercial data were still able to reduce fishing on a
protected species. \\

\added{\replaced{MixFishSim}{Our framework} allows users to explore the assumptions in modelling
	observational data and evaluate the underlying dynamics of such
	approaches at a fine spatial and temporal scale. }\added{From our
	application w}\deleted{W}e conclude that commercial data, while
containing bias, provide a useful tool for managing catches in mixed fisheries
if applied at the correct spatiotemporal scale. \\
%%
\end{abstract}

@@

Identifying appropriate spatial scales for fisheries closures \deleted{has been a
highlighted a}\added{i}s crucial to their success \citep{Costello2010, Dunn2016}.
Inference on fisheries spatial dynamics is hampered where spatial information
is coarse due to low resolution reporting of fisheries catch which is
aggregated across larger gridded areas \citep{Branch2005}. Further, if data
does not allow identification of spatial features it may lead to poorly sited
closures \replaced{that}{which} are ineffectual or have unintended consequences. For example,
increased benthic impact on previously unexploited areas from the cod closure
in the North Sea were observed without the intended effect of reducing cod
exploitation \citep{Rijnsdorp2001,Dinmore2003}).\\

More refined spatiotemporal information has since become available through the
combination of logbook and Vessel Monitoring System (VMS) data \citep{Lee2010,
	Bastardie2010, Gerritsen2012, Mateo2016} and more real-time spatial
management has been possible \citep[e.g.][]{Holmes2011}.  However, fishers
establish favoured fishing grounds through an explore-exploit strategy
\citep{Rijnsdorp2011, Bailey2018} where they search for areas with high catches
and then use experience to return to areas where they have experienced high catch
in the past. This leads to an inherently biased sampling where target species
are over-represented in the catch as fishers exploit areas of high abundance.
There is a need to understand the influence of these biases on any spatial
management measures which are implemented based on inference from commercial
landings or catch data. \\

@@

\begin{enumerate}
	\item Do different source\added{s} of sampling-derived fisheries data
		reflect\deleted{s} the underlying community structure? 	
	\item How do\deleted{es} data aggregation and data source impact on the success
		of spatial fisheries management measures?
\end{enumerate}

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time, using a Gaussian spatial
process to define suitable habitat for different populations. As the precise
locations of the fish are known directly rather than inferred from sampling
or commercial catch, we can use the population model to validate how inference
from fisheries-dependent and fisheries independent sampling relates to the real
community structure in a way we could not with real data. We ii) compare, at
different spatial and temporal aggregations, the real (simulated) population
distributions to samples from fisheries-dependent and fisheries independent \textcolor{blue}{[CM: hyphenate or don't both, i.e., fishery-independent]}
catches to test if these are a true reflection of the relative density of the
populations. We then iii) simulate a fishery closure to protect a species based
on different spatial and temporal data aggregations. \\

@@

Fishing locations were chosen based on random search and, with increasing
proportion as time progressed, experience of profitable catches built up in the
same month from previous years and from the previous trip.  `Profitable' in
this context was defined as the locations where the top 70 \% of expected
profit would be found given revenue from previous trips and cost of movement to
the new fishing location. This probability was based on a logistic sigmoid
function with a lower asymptote of 0 and upper asymptote of 0.95, and a growth
rate that ensures the upper asymptote (where decisions are mainly based on past
knowledge) is reached approximately halfway through the simulation.  \\

@@

\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_space_Rev2}
	\caption{[Colour] Data aggregation at different spatial resolutions
		over a ten year period. \added{The figure shows catch
			composition at each spatial unit represented by a
			square pie chart of the four populations. The area of
			each colour is proportionate to the weight of each
			population caught in that unit. Figure produced using
			the R package `mapplots' (\cite{Gerritsen2014}.}
	}
	\label{fig:1}
\end{figure}
